Concurrent Programming with Harmony

### 2020-07-28 book/paper.tex
Made pass over Chapter 9

It is important to appreciate the difference between an
\emph{atomic section} (the statements executed within an
\textbf{atomic} statement) and a \emph{critical section}
(protected by a lock of some sort).
The former ensures that while the
atomic statement is executing no other process can execute.
The latter allows multiple processes to run concurrently,
just not within the critical section.
The former is rarely available to a programmer, while the latter
is very common.

In Harmony, atomic statements allow you to \emph{implement} your own
synchronization primitives like test-and-set.  Other common examples
include compare-and swap and fetch-and-add.  Atomic statements
are not intended to \emph{replace} locks or other synchonization primitives.
When solving synchronization problems you should not directly use
atomic statement but use the synchronization primitives that are available
to you.  But if you want to design a new synchronization primitive, then
use \textbf{atomic} by all means.

\begin{figure}
\begin{code}
\harmonysource{lockintf}
\end{code}
\caption{The \texttt{Lock} interface in the \texttt{synch} module.}
\label{fig:spinlocks}
\end{figure}

@@

runs the file \texttt{x.hny} using the suspension version of the \texttt{synch} module.
You will find that using the \texttt{synchS} module often leads to searching a
significantly larger search space than using the \texttt{synch} module.
Part of the reason is that the \texttt{synchS} module keeps track of the order
in which processes wait for a lock.

\section*{Exercises}
\begin{problems}
\item Write a new synchronization module, similar to \autoref{fig:spinlocks}, but using
atomic swap.
\item
Run \autoref{fig:incfixed} using (1) \texttt{synch}, (2), \texttt{synchS},
and (3) your own module.  Report how many states were explored by Harmony for each
version of the module.
\item \label{ex:trylock} Implement \texttt{trylock}(\textit{lk}) as an additional
interface to the \texttt{synch} and \texttt{synchS} modules.
This interface is like \texttt{lock}(\textit{lk}) but never blocks.  It
returns \texttt{True} if the lock was available (and now acquired) or \texttt{False}
if the lock was already taken.
\end{problems}

>>>>

It is important to appreciate the difference between an
\emph{atomic section} (the statements executed within an
\textbf{atomic} statement) and a \emph{critical section}
(protected by a lock of some sort).
The former ensures that while the
\texttt{atomic} statement is executing no other process can execute.
The latter allows multiple processes to run concurrently,
just not within the critical section.
The former is rarely available to a programmer (e.g., none of
Python, C, or Java support it), while the latter
is very common.

In Harmony, \texttt{atomic} statements allow you to \emph{implement} your own
synchronization primitives like test-and-set.  Other common examples
include compare-and swap and fetch-and-add.  \texttt{atomic} statements
are not intended to \emph{replace} locks or other synchonization primitives.
When solving synchronization problems you should not directly use
\emph{atomic} statements but use the synchronization primitives that are available
to you.  But if you want to design a new synchronization primitive, then
use \textbf{atomic} by all means.

\begin{figure}
\begin{code}
\harmonysource{lockintf}
\end{code}
\caption{The \texttt{Lock} interface and implementation in the \texttt{synch} module.}
\label{fig:spinlocks}
\end{figure}

@@

runs the file \texttt{x.hny} using the suspension version of the \texttt{synch} module.
You will find that using the \texttt{synchS} module often leads to searching a
significantly larger state space than using the \texttt{synch} module.
Part of the reason is that the \texttt{synchS} module keeps track of the order
in which processes wait for a lock, while the \texttt{synch} module does not.

\section*{Exercises}
\begin{problems}
\item Write a new synchronization module, similar to \autoref{fig:spinlocks}, but using
atomic swap (similar to \autoref{ex:swap}).
\item
Run \autoref{fig:incfixed} using (1) \texttt{synch}, (2), \texttt{synchS},
and (3) your own module.  Report how many states were explored by Harmony for each
version of the module.
\item \label{ex:trylock} Implement \texttt{trylock}(\textit{lk}) as an additional
interface for both the \texttt{synch} and \texttt{synchS} modules.
This interface is like \texttt{lock}(\textit{lk}) but never blocks.  It
returns \texttt{True} if the lock was available (and now acquired) or \texttt{False}
if the lock was already taken.
\end{problems}
